Place of death in NRS COVID-19 deaths: visualisation for 2022 data

### 2022-10-27 covid_nrs_place_of_death_visualisations_2022.Rmd
completed a first draft version of the 'years skipped ahead' plot.

>>>>

# Years 'skipped ahead' due to Covid

There was a pre-pandemic trend of increasing deaths at home - by plotting a regression line we can compute the 'years skipped ahead' equivalent to the increase in deaths at home due to covid

```{r}
regression_data <- 
  merged_proportions_of_deaths_by_place %>% 
  filter(place_of_death == "Home & other non-institution" & year %in% 2015:2019) %>% 
  mutate(years_since_2015 = 0:(n()-1))

regression_model <- lm(data = regression_data, formula = number_of_deaths ~ years_since_2015)

predicted_data <- tibble(years_since_2015 = seq(from=0, to=30, by = 0.1)) %>%
  mutate(year = 2015 + years_since_2015,
         predicted = predict.lm(object = regression_model, newdata = tibble(years_since_2015))
         )

## 2022 only has X weeks' worth of data available; we can extrapolate the annual
## total by computing the percentage of annual data represented by previous
## years at X weeks, and then averaging across those, or taking the min & max
extrapolated_2020_data <-
  historical_weekly_average_overall %>%
  filter(place_of_death == "Home & other non-institution" & historic_period == "2015-2019") %>%
  summarise(
    deaths_mean_so_far = sum(deaths_mean[week_number <= weeks_available_2022], na.rm= TRUE),
    deaths_mean_total = sum(deaths_mean, na.rm= TRUE),
    proportion_of_deaths_so_far = deaths_mean_so_far / deaths_mean_total
    )

predicted_data_annotations <- tribble(
  ~x, ~y, ~label,
  2021, 20e3, "Predicted home deaths 2020-2022 using 2015-2019 data",
  2033, 23e3, "Actual home deaths 2020-2022*",
  2033, 17.5e3, "If the 2015-2019 trend continued we wouldn't observe this number of home deaths until 2031",
)

actual_and_predicted_data <-
  bind_rows(
    merged_proportions_of_deaths_by_place %>%
    select(year, number_of_deaths, place_of_death) %>%
    filter(place_of_death == "Home & other non-institution" & year %in% 2015:2019) %>%
      mutate(data_type = "Actual")
    ,
    merged_proportions_of_deaths_by_place %>%
    select(year, number_of_deaths, place_of_death) %>%
    filter(place_of_death == "Home & other non-institution" & year %in% 2020:2022) %>%
    mutate(
      data_type = "Predicted",
      number_of_deaths = if_else(year %in% 2020:2022, compute_y_from_x_using_regression(x = year-2015, regression_model = regression_model), number_of_deaths)
      )
    ,
    merged_proportions_of_deaths_by_place %>%
      select(year, number_of_deaths, place_of_death) %>%
      filter(place_of_death == "Home & other non-institution" & year %in% 2020:2022) %>%
      mutate(
        number_of_deaths = if_else(year == 2022, number_of_deaths / extrapolated_2020_data$proportion_of_deaths_so_far, number_of_deaths),
        year = 2015 + compute_x_from_y_using_regression(y = number_of_deaths, regression_model = regression_model),  # compute predicted X
        data_type = "Actual"  # note: this is actual figures but
      )
  )

fig_expected_home_deaths_regression <-
  ggplot(data = actual_and_predicted_data, aes(x=year, y=number_of_deaths)) +
  geom_rect(aes(xmin = 2019.5, xmax = 2022.5, ymin = 15.5e3, ymax = 18.5e3), fill = "yellow", alpha = 0.3) +
  geom_rect(aes(xmin = 2030, xmax = 2035, ymin = 19.5e3, ymax = 22e3), fill = "yellow", alpha = 0.3) +
  geom_text(data = predicted_data_annotations, aes(x=x,y=y,label=str_wrap(label, width = 40))) +
  geom_point(aes(shape = data_type)) +
  geom_line(data = predicted_data, aes(x = year, y = predicted), linetype = "dotted") +
  theme_minimal() +
  theme(legend.position = "top", legend.title = element_blank()) +
  xlim(c(2015,2037)) +
  scale_y_continuous(labels = scales::comma) +
  labs(y = NULL, x = "Year (hypothetical)", title = "Increasing deaths at home", subtitle = str_wrap("Home deaths were on the increase before the pandemic.\nIf we assumed the same annual increases from 2015-2019 we wouldn't expect the pandemic levels of home deaths until 2031!", width = 90), caption = paste0(c(glue::glue("*Annual home deaths for 2022 extrapolated from data for weeks 1-{weeks_available_2022}"),source_nrs_text), collapse="\n"))

fig_expected_home_deaths_regression %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_expected_home_deaths_regression", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 18, height = 12, dpi = 300, units = "cm")
```
